---
title: 'Psych 251 PS4: Simulation'
author: "Griffin Dietz"
date: "November 11, 2018"
output: 
  html_document:
    toc: true
---

This is problem set #4, in which we want you to integrate your knowledge of data wrangling with some basic simulation skills. It's a short problem set to help you get your feet wet in testing statistical concepts through "making up data" rather than consulting a textbook or doing math. 

For ease of reading, please separate your answers from our text by marking our text with the `>` character (indicating quotes). 

```{r warning=FALSE}
library(tidyverse)
```

Let's start by convincing ourselves that t-tests have the appropriate false positive rate. Run 10,000 t-tests with standard, normally-distributed data from a made up 30-person, single-measurement experiment (the command for sampling from a normal distribution is `rnorm`).

The goal of these t-tests are to determine, based on 30 observations, whether the underlying distribution (in this case a normal distribution with mean 0 and standard deviation 1) has a mean that is different from 0. In reality, the mean is not different from 0 (we sampled it using `rnorm`), but sometimes the 30 observations we get in our experiment will suggest that the mean is higher or lower. In this case, we'll get a "significant" result and incorrectly reject the null hypothesis of mean 0.

What's the proportion of "significant" results ($p < .05$) that you see?

> The mean number of significant results is about 500 (~5%).

First do this using a `for` loop.

```{r}
num_sig_for = 0
for (i in 1:10000) {
  norm_data <- rnorm(30)
  p_val <- t.test(norm_data)$p.value
  if(p_val < .05) {
    num_sig_for = num_sig_for + 1
  }
}
mean_for = num_sig_for / 10000
mean_for
```

Next, do this using the `replicate` function:

```{r}
fn <- function() {
  norm_data <- rnorm(30)
  if(t.test(norm_data)$p.value < 0.05) {
    return(1)
  }
  return(0)
}
vec <- replicate(10000, fn())
num_sig_rep = sum(vec)
mean_rep = num_sig_rep / 10000
mean_rep
```

How does this compare to the intended false-positive rate of $\alpha=0.05$?

> ANSWER

Ok, that was a bit boring. Let's try something more interesting - let's implement a p-value sniffing simulation, in the style of Simons, Nelson, & Simonsohn (2011).

Consider this scenario: you have done an experiment, again with 30 participants (one observation each, just for simplicity). The question is whether the true mean is different from 0. You aren't going to check the p-value every trial, but let's say you run 30 - then if the p-value is within the range p < .25 and p > .05, you optionally run 30 more and add those data, then test again. But if the original p value is < .05, you call it a day, and if the original is > .25, you also stop.  

First, write a function that implements this sampling regime.

```{r}
double.sample <- function (upper_p) {
  norm_data = rnorm(30)
  p_val <- t.test(norm_data)$p.value
  if (p_val < upper_p && p_val > .05) {
    norm_data <- c(norm_data,rnorm(30))
    p_val <- t.test(norm_data)$p.value
  } 
  if (p_val < .05) {
    return(1)
  }else {
    return(0)
  }
}
```

Now call this function 10k times and find out what happens. 

```{r}
vec <- replicate(10000, double.sample(.25))
num_sig_rep = sum(vec)
mean_rep = num_sig_rep / 10000
mean_rep
```

Is there an inflation of false positives? How bad is it?

> Yes. We now see the mean number of signifianct results is around 710 (~7.1%).

Now modify this code so that you can investigate this "double the sample" rule in a bit more depth. Let's see what happens when you double the sample ANY time p > .05 (not just when p < .25), or when you do it only if p < .5 or < .75. How do these choices affect the false positive rate?

HINT: Try to do this by making the function `double.sample` take the upper p value as an argument, so that you can pass this through dplyr.

HINT 2: You may need more samples. Find out by looking at how the results change from run to run.

```{r}
sum_50 <- sum(replicate(30000, double.sample(.5)))
sum_50
mean_50 <- sum_50 / 30000
mean_50

sum_75 <- sum(replicate(30000, double.sample(.75)))
sum_75
mean_75 <- sum_75 / 30000
mean_75

sum_100 <- sum(replicate(50000, double.sample(1)))
sum_100
mean_100 <- sum_100 / 50000
mean_100
```

What do you conclude on the basis of this simulation? How bad is this kind of data-dependent policy?

> When p-values are trending toward significance, adding additional data has a large capacity to affect the results. By doubling the amount of data when p-values are less than .25, we nearly are 50% more likely to find significant results in our t-test. However, as results become further from significant, adding this additional data has decreasing effect. When we double the data for all p-values, for instance, we only see an average of about 8.4% significant results, not a proportional increase over 7.1% for p-values less than .25. This is because continuing to add data when p-values are far from significant is less likely to swing the p-value to significant than adding that data when the p-value is already trending in the "right" direction.
In summary, this data-dependent policy is bad regardless of what upper p-value threshold we use, because we will end up counting more results as significant than we should. However, it has a decreasing effect on results as upper p-values increase. It highlights the importance of establishing sample size before beginning to run a study.








